---
output: pdf_document
---
## Coursera -Resgreesion Models Project

## Problem summary
You work for Motor Trend, a magazine about the automobile industry. Looking at a data set of a collection of cars, they are interested in exploring the relationship between a set of variables and miles per gallon (MPG) (outcome). They are particularly interested in the following two questions:   
The main aims of this analysis are:  

**1.** Determine which if better for MPG, automatic or manual transmission.  
**2.** Quantify the MPG difference between automatic and manual transmissions.

## 1.Determine which if better for MPG, automatic or manual transmission. 

As the first attempt to explore the data,
```{r}
## average mpg of all automatic transmission typed(am=0) cars
avg_a = mean( mtcars[ mtcars$am == 0, "mpg" ] )
avg_a

## avarge mpg of all manual transmission typed(am=1) cars
avg_m = mean( mtcars[ mtcars$am == 1, "mpg" ] )
avg_m
```
This shows at average cars with manual transmission have `r avg_m - avg_a` mileage per gallon more than those with automatic transmission. 

## 2.Quantify the MPG difference between automatic and manual transmissions

The linear regression model, described in later sections of this work, can quantify the difference more closely.


## Clean Data

Those variables with integer values are better as factor formatted.

```{r}
mydata <- mtcars

mydata$cyl <- as.factor(mydata$cyl)
mydata$vs <- as.factor(mydata$vs)
mydata$gear <- as.factor(mydata$gear)
mydata$carb <- as.factor(mydata$carb)
mydata$am <- as.factor(mydata$am)
```

And the "am" variable is assigned with meaningful labels
```{r}
levels(mydata$am) <- c('Automatic','Manual')
```

### T-test
Applying T-Test shows clearly that the manual and automatic are statistically significant different.
"Null Hypothesis": true difference in two group "manual" and "automatic" equals to 0
```{r}
t_res <- t.test(mpg ~ am, data=mydata)
t_res$p.value
```
The P-value is so small that the "Null Hypothesis" will be rejected.

## Linear Regression Model

### Selection 

```{r results = 'hide'}
model.all <- lm(mpg ~ ., data = mydata)
n <- nrow(mydata)
## Stepwise Algorithm
model.new <- step(model.all, direction = "both", k = log(n))
```

```{r}
summary(model.new)$coefficients
```

The result suggests the combinations of 3 predictors: "wt", "qsec" and "am". See the Figure 2.

Now compare the difference among models more with "anova" method.

```{r}
anova <- anova(lm(mpg ~ am, data = mydata), 
               lm(mpg ~ am + wt, data = mydata),
               model.new )

cbind(anova[1], anova[2], anova[3], anova[4], anova[5], anova[6])
```

It shows that by including new variables "wt" and "qsec" into the "lm" formula, the fitness of regression model gets improved because the Residuals Sum of Square (RSS) decreases steadily. 

And fitted model parameters
```{r, echo=FALSE}
model.new

```
indicate that, regard to the predicated mileage, the coefficient on "wt -- Weight(1000/lb)" with value `r model.new$coefficients[2]`  has even larger impact than the coefficient on "am" with value `r model.new$coefficients[4]`. With other words, the "transmission type" is not be the most decisive factor for the purpose of saving fuel cost.

### Residuals

Based on the Figure 3 in the Appendix

- The Residuals vs Fitted plot seem to verify the independence assumption as the points are randomly scattered on the plot.

- The Residuals and Leverage plot indicates no leverage detected as all points fall inside the bounds.

### Diagnostics
Find out those highly influential "cars" in the "mtcars" datasets.

```{r}
influentialPoints <- apply( influence.measures(model.new)$is.inf, 1, sum)>0
row.names(mydata[influentialPoints,])
```
## Conclusion
To the end, after developing a linear regression model, the question about the difference of mpg(outcome) due to "transmission type" has its clear answer. One **uncertainty**, look at the Figure 2, almost all automatic cars have light weight than cars with manual transmission so the original dataset is not well unbiased to study the "mpg". 


## Appendix


**Figure 1. mpg vs transimission type**

````{r, echo=FALSE, fig.width=6,fig.height=4}
boxplot(mpg ~ am, data = mydata)
title(xlab = "automatic or manual (am)", ylab = "mileage per gallon (mpg)")

```


**Figure 2. mpg vs. wt by transmission type**

```{r, echo=FALSE, fig.width=6,fig.height=4}
plot(mydata$wt, mydata$mpg, col=mydata$am, pch =18, xlab = "weight", ylab = "mpg")
legend("topright", c("Automatic", "Manual"), col=c(1,2), pch =18)
```


**Figure 3. Model Analysis**

```{r, echo=FALSE, fig.width=9,fig.height=9}
par(mfrow = c(2, 2), oma=c(0,0,2,0))
plot(model.new)
```